Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2018-05-26 JCP/manuscript.tex
JCP Communication directory with the manuscript is added

%\documentclass[aps,prl,reprint,amsmath,amssymb]{revtex4-1}
%\documentclass[10pt,aps,prl,twocolumn,amsmath,amssymb,superscriptaddress,longbibliography]{revtex4-1}
%\documentclass[preprint,showpacs,preprintnumbers,amsmath,amssymb]{revtex4-1}
%\documentclass[twocolumn,showpacs,preprintnumbers,amsmath,amssymb]{revtex4}
% Some other (several out of many) possibilities
%\documentclass[preprint,aps]{revtex4}
%\documentclass[preprint,aps,draft]{revtex4}
%\documentclass[prb,amsmath,amssymb]{revtex4}% Physical Review B

% define variable that compiles the main text (as opposed to SI text)
\newcommand*{\MAINTEXT}{}

%%% JPCL
\documentclass[journal=jpclcd,manuscript=letter]{achemso}
\usepackage{achemso}
%\usepackage[T1]{fontenc} % Use modern font encodings
%%% end JPCL

@@

\title{
\ifdefined\MAINTEXT
\else
Supplemental Material: \\
\fi
Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
}

@@

\ifdefined\MAINTEXT
\else
	\clearpage
	%\widetext
	\setcounter{figure}{0}
	\setcounter{page}{1}
	\renewcommand{\thefigure}{S\arabic{figure}}
\fi

@@

\ifdefined\MAINTEXT

\begin{tocentry}
%\begin{figure}
\includegraphics[width=2in]{TOCG.eps}
%\end{figure}
\end{tocentry}

@@

\fi

@@

%\ifdefined\MAINTEXT

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}) -- compact orbitals first described in Ref.~\citenum{a:stoll}. 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However, the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}.

@@

In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) averaged over atoms with different $m_i$
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \frac{1}{3} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
In practice, we found (see results below) that this approach is not particularly accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the ACF integral can provide a reasonable starting value of $\Delta$. This value can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.

The inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
Nevertheless, it is capable to reproduce correct dynamical properties of a system as long as $\gamma$ is set to a small value and  partially optimized ALMOs remain close to the ground state resulting in $\Delta \ll \gamma$. 
%With $\gamma$ set to zero, the nuclei will follow time-reversible Newtonian trajectories closely enough to reproduce dynamical properties of the ensemble as long as partially optimized ALMOs remain close to the ground state and $\Delta$ is small.

@@

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=13cm]{Dynamical_Data_Tiled.eps}
\caption{\label{fig:dynproperties} 
Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) RDF, 
(b) kinetic energy distribution (the gray curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) IR spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

ALMO AIMD was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
%RZK: cite our own water paper
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation. We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:pbe,Grimme2010}. 
Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
%Although not a required component of ALMO AIMD, 
The predictor of the Kolafa scheme~\cite{Kolafa2003} was adopted to localized orbitals~\cite{a:2ndcpmd} to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps of the robust two-stage optimization procedure. 
%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.

@@

\begin{figure}
\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=13cm]{DeltaForceComparison_with_ACF.eps}
\caption{\label{fig:randomforce} 
The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line. 
%(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., 
(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
(c) Normalized ACF $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it}$ of the instantaneous error in panel (a).
%Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
}
\end{figure}

Even with $\epsilon_{\text{SCF}} = 10^{-2}$, the simulation is stable with the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b). $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$. 
We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
%RZK addition for a longer article: In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
%
Analysis of $\delta \vec{f_{i}}(t)$ shows that the error indeed resembles Gaussian white noise. The mean of the error is zero (black line in Figure~\ref{fig:randomforce}a). Its ACF decays rapidly (Figure~\ref{fig:randomforce}c) so that the errors can be considered uncorrelated on time scale of $\SI{50}{\fs}$. Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water. We established that the main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the oscillating component of $\delta f$ (Figure~\ref{fig:randomforce}b). We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ 
%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
(Figure~S1 in the Supplemental Material).

@@

% RZK longer article: averaging is done over 10 time steps 
To demonstrate the computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
%In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}. 
It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions---a particularly challenging case for DM-based LS methods.
ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$, even for medium-size systems. While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today.

@@

\begin{figure}
\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=13cm]{strongscaling_log.eps}
\caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
%Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
%Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.~\cite{a:ot,a:ot2,a:2ndcpmd}
For ALMO methods, $R_{c} = 1.6$ vdWR. 
Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
}
\end{figure}

@@

\begin{figure}
\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=13cm]{weakscaling.eps}
\caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
Dashed gray lines connect systems simulated on the same number of cores to confirm LS behavior. %over a range of configurations.
}
\end{figure}

@@

% systematically improvable approximation to DFT-based AIMD.
To summarize, we demonstrated---for the first time---that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM. %for weakly-interacting systems. 
High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces. 
By obviating the optimization of the DM, the method remains remarkably efficient even with large localized basis sets. 
%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible length scales. 
The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.

\begin{acknowledgement}

The authors are grateful to Thomas K\"uhne for insightful discussions. The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059). The authors are grateful to Compute Canada and McGill HPC Centre for computer time.

\end{acknowledgement}

\bibliography{almo-langevin}
%\input{almo-langevin.ref}

%\else

\begin{suppinfo}

\subsection{Dependence of the forces on the electron localization radius}

The ALMO forces converge to the reference forces as the electron localization radius increases (Figure~\ref{fig:forcecomp}).

\begin{figure}[h!]
\includegraphics[trim={0cm 0cm 0.1cm 0.1cm},clip,width=13cm]{DeltaForceComparison_ALMO_SCF.eps}
\caption{\label{fig:forcecomp} Dependence of the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$ (red line) on the ALMO localization radius $R_c$ expressed in units of atomic van der Waals radii. Black line is the magnitude of the time average of the instantaneous error vector. %calculated as the average difference between fully converged Born-Oppenheimer forces and fully converged ALMO DFT forces, averaged over all dimensions and all molecules for each time step. 
Forces are computed with the PBE XC functional, TZV2P basis set, and $\epsilon_{\text{SCF}} = 10^{-7}$~a.u. for the configurations from a 30-ps trajectory generated with delocalized AIMD at T=298~K.}
\end{figure}

%\begin{figure}[!htb]
%\includegraphics[trim={1.7cm 0.5cm 3.3cm 0.1cm},clip,width=\textwidth]{strongscaling_linear.eps}
%\caption{\label{fig:strongscaling_linear} Linear plots showing LS behaviour of all ALMO methods for liquid water at 298 K. 
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, averaged over 10 MD steps. 
%The localization radius was set to $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u.}
%\end{figure}

%\begin{figure}[!htb]
%\includegraphics[trim={2.4cm 0cm 3.0cm 1cm},clip,width=8.6cm]{Diffusion_Comparison.eps}
%\caption{\label{fig:diffusion} Calculated mean-squared displacement function of water using ALMO DFT with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR.
%Compared with equivalent MD run using OT DFT as a reference.
%Calculations were performed at 298 K with the dispersion-corrected PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 30 ps.}
%\end{figure}

\end{suppinfo}

%\fi % SI - main text switch

>>>>

\documentclass[aip,jcp,reprint,amsmath,amssymb]{revtex4-1}

@@

\title{
Compact orbitals enable low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems
}

@@

@@

%\begin{tocentry}
%%\begin{figure}
%\includegraphics[width=2in]{TOCG.eps}
%%\end{figure}
%\end{tocentry}

@@

@@

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}) -- compact orbitals first described in Ref.~\citenum{a:stoll}. 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However, the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}. ALMO DFT agorithms are presented in the Supporting Information.

@@

In principle, $\Delta$ can be calculated using the integral of Eq.~(\ref{eq:stochastic3}) averaged over atoms with different $m_i$
%
\begin{align}
\label{eq:delta}
\Delta &= (2 k_B T m_i )^{-1} \int_{-\infty}^{\infty} \frac{1}{3} \langle \delta \vec{f}_i (0) \cdot \delta\vec{f}_i(\tau) \rangle d\tau
\end{align}
%
if one can afford computing the reference forces $f^{\text{KS}}_{i\alpha}(t)$ (i.e. $R_c \rightarrow \infty$ and $\epsilon_{\text{SCF}} \rightarrow 0$) for a short representative AIMD trajectory. 
In practice, we found (see results below) that this approach is not sufficiently accurate because the $\delta_{ij}\delta_{\alpha\beta}$ assumption in Eq.~(\ref{eq:stochastic3}) does not strictly hold. 
Nevertheless, the integral of the autocorrelation function (ACF) in Eq.~\ref{eq:delta} can provide a reasonable starting value of $\Delta$. %If the reference forces cannot be computed even for a short simulations, then a starting calue of $\Delta$ can be determined . 
The estimated $\Delta$ can be further fine-tuned in a series of short trial-and-error ALMO AIMD runs until the average kinetic energy corresponds to the requested temperature $\langle \frac{1}{2} m_i \dot{\bm{r}}^{2}_{i} \rangle = \frac{3}{2} k_{B} T$.

The inherently stochastic approach presented here does not aim to produce fully time-reversible dynamics for atomic nuclei. 
Nevertheless, it is capable to reproduce correct dynamical properties of a system as long as $\gamma$ is set to a small value and  partially optimized ALMOs remain close to the ground state resulting in $\Delta \ll \gamma$. 
An alternative approach to performing microcanonical MO-based LS AIMD would be to combine ALMO DFT with the extended Lagrangian method described in Ref.~\onlinecite{Niklasson2008} that has been shown to conserve the total energy in DM-based simulations~\cite{Cawkwell2012}.

% RZK
% Niklasson2008  --> DOI: 10.1103/PhysRevLett.100.123004
% Cawkwell2012 -->  doi: 10.1063/1.4755991

@@

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{1.eps}
\caption{\label{fig:dynproperties} 
Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) RDF, 
(b) kinetic energy distribution (the gray curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) IR spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

ALMO AIMD was implemented in CP2K, an open source materials modeling package~\cite{www:cp2k}. 
Accuracy and efficiency of ALMO AIMD was tested using liquid water as an example. 
%RZK: cite our own water paper
This system is challenging because intermolecular electron delocalization is a critical component of hydrogen bonding and must be described correctly to reproduce static and dynamical properties of liquid water. 
%To calculate properties of water at ambient conditions, 
A periodic cell containing 125 molecules was simulated for $\SI{30}{\ps}$ at $T=\SI{298}{\K}$ and a constant density of $\SI{1.01}{\g\cdot\cm^{-3}}$. 
The initial configurations of the water boxes were created by equilibrating the system for 20~ps using the fully converged orbitals. 
Ricci-Ciccotti algorithm~\cite{Ricci2003} was used to integrate the Langevin equation (see Supporting Information). We found that $\gamma = \SI{e-3}{\per\fs}$ is large enough to thermostat the system efficiently and small enough not to significantly affect dynamical properties of liquid water.
%
In the dual Gaussian and plane-wave scheme implemented in CP2K~\cite{a:quickstep}, the TZV2P basis set was used to represent molecular orbitals, and a plane-wave cutoff of $\SI{320}{Ry}$ used to represent electron density. 
The XC energy was approximated using the dispersion-corrected PBE functional~\cite{a:pbe,Grimme2010}. 
Separable norm-conserving pseudopotentials were used~\cite{a:hgh} and the Brillouin zone was sampled at the $\Gamma$-point. 
%Although not a required component of ALMO AIMD, 
The predictor of the Kolafa scheme~\cite{Kolafa2003} was adopted to localized orbitals~\cite{a:2ndcpmd} to generate a highly accurate initial ALMOs in both SCF stages, which can be brought close to the ground state with just a few SCF steps of the robust two-stage optimization procedure. 
%RZK: The time-reversibility of ASPC is not a critical reqirement of the proposed method.

@@

\begin{figure}
\includegraphics[trim={0.5cm 0cm 0.7cm 0.1cm},clip,width=8.6cm]{2.eps}
\caption{\label{fig:randomforce} 
The red line is the Euclidean norm of the instantaneous error $\langle \| \delta \vec{f_{i}}(t) \| \rangle_{i}$, the black line is the magnitude of the time average of the instantaneous error vector, and the green line is the time average of the red line. 
%(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. relative to the fully converged ALMO reference, 
(a) $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u., 
(b) $R_{c} = 1.6$ vdWR and fully converged ALMO SCF, 
(c) Normalized ACF $\frac{1}{3}\langle \delta \vec{f}_i (t) \cdot \delta\vec{f}_i(t+\tau) \rangle_{it}$ of the instantaneous error in panel (a).
%Forces are computed with the PBE XC functional and TZV2P basis set for the configurations from a 10-ps trajectory generated with KS-based AIMD at T=298~K. 
}
\end{figure}

Even with $\epsilon_{\text{SCF}} = 10^{-2}$, the simulation is stable with the correct average temperature and perfect Maxwell-Boltzmann distribution (Figure~\ref{fig:dynproperties}b). $\Delta$ was initially estimated at $\SI{2e-5}{\per\fs}$ using Eq.~(\ref{eq:delta}) and then refined heuristically to $\SI{6e-5}{\per\fs}$. 
We found that it is easier to optimize $\Delta$ when $\gamma$ is set to zero because of reduced noise in the trial runs. 
%RZK addition for a longer article: In theory, $\Delta$ should be independent on system size. However, we found that $\Delta$ dependents on system size due to finite size effects, which decay for increasing system size.
%
Analysis of $\delta \vec{f_{i}}(t)$ shows that the error indeed resembles Gaussian white noise. The mean of the error is zero (black line in Figure~\ref{fig:randomforce}a). Its ACF decays rapidly (Figure~\ref{fig:randomforce}c) so that the errors can be considered uncorrelated on time scale of $\SI{50}{\fs}$. Thus the main assumption behind our approach to ALMO AIMD is justified for liquid water. We established that the main source of error in forces for this system is the loose convergence criterion and not the finite $R_c$: fully converged ALMO SCF calculations remove the oscillating component of $\delta f$ (Figure~\ref{fig:randomforce}b). We also verified that the ALMO forces converge to the reference forces in the limit $R_{c} \rightarrow \infty$ 
%(Figure~\ref{fig:forcecomp} in the Supplemental Material).
(Figure~SZZZ in the Supporting Information).

@@

% RZK longer article: averaging is done over 10 time steps 
To demonstrate the computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
%In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}. 
It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions---a particularly challenging case for DM-based LS methods.
ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$, even for medium-size systems. While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today. For comparison, the crossover point between the cubically-scaling DFT and a representative LS method that computes a sparse DM as the matrix sign function of the effective Hamiltonian~\cite{a:ls-dm-sign} is expected to lie well above 10,000 water molecules (Figures 5a and S5 in Ref.~\onlinecite{:a:almo-ls}).

@@

\begin{figure}
\includegraphics[trim={2.5cm 0.5cm 3.4cm 0.9cm},clip,width=8.6cm]{3.eps}
\caption{\label{fig:strongscaling_log} Timing benchmarks for PBE/TZV2P simulations of liquid water on 256 compute cores. 
%Logscale plots which show the asymptotic LS behaviour of ALMO methods for liquid water at 298 K.
%Compared with the orbital transformation using fully converged SCF and second generation Car-Parinello force calculation methods.~\cite{a:ot,a:ot2,a:2ndcpmd}
For ALMO methods, $R_{c} = 1.6$ vdWR. 
Cyan lines represent perfect cubic scaling, whereas gray lines represent perfect linear scaling. 
}
\end{figure}

@@

\begin{figure}
\includegraphics[trim={1.6cm 0.8cm 4.7cm 0cm},clip,width=8.6cm]{4.eps}
\caption{\label{fig:weakscaling} Weak scalability benchmarks for PBE/TZV2P ALMO AIMD with  $R_{c} = 1.6$ vdWR and $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. 
Dashed gray lines connect systems simulated on the same number of cores to confirm LS behavior. %over a range of configurations.
}
\end{figure}

@@

% systematically improvable approximation to DFT-based AIMD.
To summarize, we demonstrated---for the first time---that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM. %for weakly-interacting systems. 
High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces. 
By obviating the optimization of the DM, the method remains remarkably efficient even with large localized basis sets. 
%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible length scales. 
The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
A generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.

%\begin{suppinfo}

%Convergence of approximate forces to the reference forces with the electron localization radius.

%\end{suppinfo}

%\begin{acknowledgement}

\textbf{Acknowledgements.} The authors are grateful to Thomas K\"uhne for insightful discussions. The research was funded by the Natural Sciences and Engineering Research Council of Canada through the Discovery Grant (RGPIN-2016-05059). The authors are grateful to Compute Canada and McGill HPC Centre for computer time.

%\end{acknowledgement}

\bibliography{manuscript}
